### 2024-01-28 Evalsims.Rmd
sims

---
title: "Untitled"
output: html_document
date: '2024-01-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Maybe use the two plots of the blocks tomorrow, with RF as less calibration and GAM as mire

```{r}
library(data.table)
library(ggplot2)
out <- rbindlist(lapply(c(1, 1.5, 2, 2.5, 3), function(shape) {
  data <- fread(paste0("results/sims_calibration_1000_rf_5_0.1_", shape, "_0.6.csv"))
  data$shape <- shape
  data
}))

method_names <- c("HistCAL-CP (5 bins)", "Conditional-CP", "SC-CP (ours)", "Unconditional-CP", "Mondrian-CP (5 bins)", "Oracle")
names(method_names) <- c("cal_binning_1", "conditional", "isotonic","marginal", "mondrian_1", "oracle")

out <- out[out$method %in% c("cal_binning_1", "mondrian_1", "marginal", "isotonic", "conditional", "oracle")]

out$method <- method_names[match(out$method, names(method_names))]

out$method <- factor(out$method, levels = 
                     c("SC-CP (ours)", "Conditional-CP" , "Unconditional-CP", "HistCAL-CP (5 bins)", "Mondrian-CP (5 bins)", "Oracle")  )

plt1 <- ggplot(out[out$bin == "marginal"], aes(x = shape, y = width, color = method)) + geom_point(aes(shape = method), size = 3)   + theme_bw() + labs(x = TeX("Distribution shift in training data ($\\kappa_{train}$)"), y = "Interval Width") +   guides(color = guide_legend(title = element_blank()), shape = guide_legend(title = element_blank())) + 
 theme(legend.text = element_text(size = 11), legend.key.size = unit(2, "lines"))

plt1
ggsave("plots/calibration_shift.pdf" , width = 7, height = 5)

plt2 <- ggplot(out[out$bin == "marginal"], aes(x = shape, y = coverage, color = method)) + geom_point(aes(shape = method), size = 3)   + theme_bw() + labs(x = TeX("Distribution shift in training data ($\\kappa_{train}$)"), y = TeX("Coverage ($\\alpha = 0.1$)")) +   guides(color = guide_legend(title = element_blank()), shape = guide_legend(title = element_blank())) + 
 theme(legend.text = element_text(size = 11), legend.key.size = unit(2, "lines"))

plt2

#combined_plot <- plt1 + plt2 + plot_layout(guides = "collect") & theme(legend.position = "bottom")

#ggsave("plots/calibration_shift.pdf" , width = 7, height = 5)

```

```{r}
library(data.table)
library(ggplot2)
library(latex2exp)
method_names <- c("HistCAL (5 bins)", "Conditional-CP", "SC-CP (ours)", "Unconditional-CP", "Mondrian-CP (5 bins)", "Oracle")
names(method_names) <- c("cal_binning_5", "conditional", "isotonic","marginal", "mondrian_5", "oracle")

out <- fread(paste0("results/sims_coverage_1000_gam_5_0.1_1_0.6.csv"))
#out <- fread(paste0("results/sims_calibration_1000_rf_5_0.1_1_0.6.csv"))

out <- out[out$method %in% c("oracle", "cal_binning_5", "mondrian_5", "marginal", "isotonic", "conditional")]
out$method <- method_names[match(out$method, names(method_names))]

out$method <- factor(out$method, levels = 
                     c("SC-CP (ours)", "Conditional-CP" , "Unconditional-CP", "HistCAL (5 bins)", "Mondrian-CP (5 bins)", "Oracle")  )

plot1 <- ggplot(out , aes(x = bin, y = coverage, color = method)) + geom_point(aes(shape = method), size = 3) + theme_bw() + labs(y = TeX("Coverage ($\\alpha = 0.1$)"), x = TeX("Quintiles of $\\sigma^2(X)$")) + scale_x_discrete(labels = c("Q.1", 
                                                       "Q.2",
                                                       "Q.3",
                                                       "Q.4",
                                                       "Q.5",
                                                       "Uncond."))  +   guides(color = guide_legend(title = element_blank()), shape = guide_legend(title = element_blank()))

plot2 <- ggplot(out , aes(x = bin, y = width, color = method)) + geom_point(aes(shape = method), size = 3)   + theme_bw() + labs(y = "Interval Width", x = TeX("Quintiles of $\\sigma^2(X)$")) + scale_x_discrete(labels = c("Q.1", 
                                                       "Q.2",
                                                       "Q.3",
                                                       "Q.4",
                                                       "Q.5",
                                                       "Uncond."))   +   guides(color = guide_legend(title = element_blank()), shape = guide_legend(title = element_blank()))

library(patchwork)
combined_plot <- plot1 + plot2 + plot_layout(guides = "collect") & theme(legend.position = "")
combined_plot
print(combined_plot)

ggsave("plots/conditional_coverage.pdf", plot = combined_plot, width = 6, height = 3.5)
# this plot is good

plot1 + plot2 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```